---
title: "Notes and Commands"
author: "Mary Kendig"
date: "February 9, 2016"
output: pdf_document
---


```{r}
5+5
lm()

??sum()

getwd()

mydata <- read.csv("C:/Users/Mary/Documents/Other Files/R/Try This/Practice R Space/r_programming_bsos_winter2016 WORKING FILE/Notes/mappingdata.csv")







??file.choose()

```




```{r}
#all the packages you should download!!!

install.packages("graphics")
library(graphics)
install.packages("NLP")
library(NLP)
install.packages("scatterplot3d")
library(scatterplot3d)
install.packages("tm")
library(tm)
install.packages("tidyr")
library(tidyr)
install.packages("dplyr")
library(dplyr)
install.packages("reshape2")
library(reshape2)
install.packages("ggplot")
library(ggplot)
install.packages("ggplot2")
library(ggplot2)
install.packages("plotly")
library(plotly)
install.packages("readxl")
library(readxl)
install.packages("xlsx")
library(xlsx)
install.packages("DTK")
library(DTK)
install.packages("MASS")
library(MASS)
install.packages("plotGoogleMaps")
library(plotGoogleMaps)
install.packages("ggmap")
library(ggmap)
install.packages("mapproj")
library(mapproj)
install.packages("scales")
library(scales)
install.packages("ggthemes")
library(ggthemes)
install.packages("extrafont")
library(extrafont)
install.packages("stringr")
library(stringr)
install.packages("vcd")
library(vcd)

data()

AirPassengers 

```

```{r}
Day #1

objects() # display objects in workspace
ls()      # same as objects()
rm("x")   # remove object from workspace
??
help()
help.search()
install.packages("x")
library(x)
ls()
load("x.RData")
save.image()
save.image("myworkspace.RData")
q("yes") # quit and save workspace as .RData
q("no")  # what I always do
getwd()  # get working directory
setwd()  # set working directory 
data()   # load R data 
head()   # first 5 rows
tail()   # last 5 rows
length() # number of columns 
nrow()   # number of rows
head(sleep, 10) # 10 rows of sleep 
summary()
summary.data.fram()

# Data

mean()
min()
max()
median()
mode()
summary(mydata) 
class(mydata$Year_level)

mydata$Year_level <- as.character(mydata$Year_level)



# Creating Vectors 
c(1, 3, 4, 2)  # ORDER MATTERS
c(1, "pie") # coersion happens
c(1, pi)    # both numbers!

# Combining Vectors
y <- c(1, 2, 3, pi)
z <- c(3, 6, 23)
c(y, z)


#Matrix: two-dimensional vector 
matrix(1:25, ncol = 5) # ncol = 5 tells how many columns 

# Array
array(1:27, dim = c(3, 3, 3))

# List
list(1, "pie", pi, x)

# Subsetting - looking at a specific section
# use [ ] or [[ ]]
yz
yz[1]  # looks at the number in the first column
yz[c(4, 3)] # fourth number and the third number in the column
mydata


# specificy specific rows and columns
# data[ , ] == data  (all rows, all columns)

data[ , ] # data[ROWS, COLUMNS]
data[1:5, ] # specify first 5 rows, and all of the columns
data[ , c(3, 1)] # specify all rows, then column 3 and column 1

# you can also call up a column by its name 
colnames(data)
data[1:5, "column1"]
data[1:5, c("column1","column2")]
colnames(data)[3] # use this function to call up specific column
colnames(data)[3] <-"subject" # use this function to rename a column

```



```{r}
```{r}
# p - value
pnorm(-8.94)*2 #for the two tailed test 

# pt(zscore, degrees of freedom)
# degrees of freedom = N - 1 
pt(2.75, 24, lower.tail = FALSE)*2

```

```{r}
#z score
pop_sd <- sd(x)*sqrt((length(x)-1)/(length(x)))
pop_mean <- mean(x)
#any x value, chose 72
z <- ((72 - pop_mean)/pop_sd)

#yellow left # blue right
yellowp1<- pnorm(72, pop_mean, pop_sd)
#or
yellowp2 <- pnorm(z)

#blue
blue <- 1 - yellowp1
blue <- 1 - yellowp2
```

```{r}
#t.test()

levels(mydata2$genre)

mydata_metal <- mydata2[mydata2$genre == "metal", ]
mydata_folk <- mydata2[mydata2$genre == "folk", ]

mydata <- rbind(mydata_metal, mydata_folk)

mydata
mydata_metal
mydata_folk

summary(mydata_metal)
summary(mydata_folk)

mean(mydata_metal$loudness)
mean(mydata_folk$loudness)

sd(mydata_metal$loudness)
sd(mydata_folk$loudness)

t.test(mydata_metal$loudness, mydata_folk$loudness)

```

```{r}
#Chisquare
# pschisq(x^2, df, lower.tail = FALSE)

pschisq(7.2, 4, lower.tail = FALSE)

#Chisquare test of independence 

```

```{r}
## test overall model and significance of each IV
summary(lm(Y~x1+x2+x3,data=d))

## compare model with student socio-demographics with model that also includes studying behavior
m1 = lm(Y~x1+x2+x3, data=d)
m2 = lm(Y~x1+x2, data=d)

anova(m1,m2)

## save best model
mbest = lm(Y~x1+x2+x3,data=d)
summary(mbest)

# save Yhat = predicted = pred and
# residuals to investigate assumptions
pred = mbest$fitted.values
resid = mbest$residuals

#############  CAR data set  ########

## read in data and ignore missing values
d = read.csv("Cars.csv")
d1 = d[order(d$engine),]
d1 = subset(d1,!is.na(engine)&!is.na(mpg))

## run regression, save predicted values and residuals
m = lm(mpg~engine,data=d1)
summary(m)
pred = m$fitted.values
resid = m$residuals

## non-linearity
## scatterplot of Xs and Y
plot(d1$engine,d1$mpg)

## outliers
## are there any unusually large residuals
sdresid = sd(resid)
resid[abs(resid)>sdresid*3]
d1[row.names(d1) %in% c(35,330,333,337,403,396),] 

## multicollinearity
## only one IV for this problem, but compute correlations between ivs
## e.g. cor.test(d1$engine,d1$horse) if we also included horse power as an IV

## test of independence
## data is not timeseries so, we don't need to worry about this as much

## constant error
## look at residuals across predicted values
plot(pred,resid)

## normality of errors
qqnorm(resid)
qqline(resid)
shapiro.test(resid)


```

```{r}
# Interaction

setwd("/Users/Yla/Documents/Teaching/Fall2015/Lecture12")

## Helper Libraries & Code
library(ggplot2) # loads graphic functions from ggplot library
# from http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/
# cookbook-r has many useful code snippets to get you started
source("HelperCode.R") # loads code with extra functions


######  EXAMPLE 0 #####
d0 = read.csv("Eysenck.csv") ## read in file
d0$Condition = factor(d0$Condition,levels=c("Counting","Rhyming","Adjective","Imagery","Intentional")) # order levels so that they make sense conceptually

head(d0) ## check that file was read in properly, examine file


### ANOVA - WITH INTERACTION ####
## ANOVA way
m = aov(Recall~Age*Condition,data=d0)
summary(m)

write.csv(mydata, file = "test.csv")

####  PLOT BARGRAPH - WITH SE ####

# calculate sd, se, 95% confidence intervals
d0.ci = summarySE(d0,measurevar="Recall",groupvars=c("Condition","Age"))

d0.ci # print new data frame and notice that it has been summarized

## create plot se
ggplot(d0.ci,aes(x=Condition,y=Recall,fill=Age))+
geom_bar(position=position_dodge(),stat="identity")+
geom_errorbar(aes(ymin=Recall-se,ymax=Recall+se),
					width=.2,
					position=position_dodge(.9))

#### Pairwise Comparisons ####


### age - counting
counting = subset(d0,Condition=="Counting")
summary(aov(Recall~Age,data=counting))

### age - rhyming
rhyming = subset(d0,Condition=="Rhyming")
summary(aov(Recall~Age,data=rhyming))

### age - adjective
adjective = subset(d0,Condition=="Adjective")
summary(aov(Recall~Age,data=adjective))

### age - imagery
imagery = subset(d0,Condition=="Imagery")
summary(aov(Recall~Age,data=imagery))

### age - intentional
intentional = subset(d0,Condition=="Intentional")
summary(aov(Recall~Age,data=intentional))



## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)


summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}

```






```



